---
title: "IDG-DREAM Round 2 Bayes Factor Analysis - all predictions"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

Round 2 of the IDG-DREAM Challenge gave participants two opportunities to predict 394 Kd values between 25 compounds and 207 kinases. There are two subchallenges in this round that participants can win via:

Sub-challenge 1: The top-performing teams will be selected based on the bootstrapped Spearman correlation of the predictions to the test dataset. Teams that have a Bayes Factor of less than 3 are tied. The tiebreaking metric to select the top performer will be the submission with the highest average AUC score.

Sub-challenge 2: The top-performing teams will be selected based on the bootstrapped RMSE of the predictions to the test dataset. Teams that have a Bayes Factor of less than 3 are tied. The tiebreaking metric to select the top performer will be the submission with the highest average AUC score.

In addition, participants/teams needed to submit a Dockerized model, writeup, and methods survey to qualified. Participants that could not fulfill this criteria were disqualified.

Here, we are only looking all results regardles of qualification.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(challengescoring)
library(reticulate)
library(tidyverse)
library(doMC)
doMC::registerDoMC(cores = detectCores()-1)

use_python("/usr/local/bin/python2")
synapse <- import("synapseclient")
syn <- synapse$Synapse()
synutils <- synapse$utils
syn$login()
source_python('https://raw.githubusercontent.com/Sage-Bionetworks/IDG-DREAM-Challenge-Analysis/master/round_1b/evaluation_metrics_python2.py?token=AE3WNSHKS34O3U47KUCXXD25ALMY2')

spearman_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  spearman(gold_py, pred_py)
}

rmse_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  rmse(gold_py, pred_py)
}

computeBayesFactorWhereRefIsNotBest <- function(bootstrapMetricMatrix,
                               refPredIndex,
                               invertBayes){

    M <- as.data.frame(bootstrapMetricMatrix - bootstrapMetricMatrix[,refPredIndex])
    K <- apply(M ,2, function(x) {
      k <- sum(x >= 0)/sum(x < 0)
      if(sum(x >= 0) > sum(x < 0)){
      return(k)
      }else{
      return(1/k)
      }
    })
    if(invertBayes == T){K <- 1/K}
    K[refPredIndex] <- 0

    return(K)
}

fv <- syn$tableQuery("select id, submissionId AS objectId, teamId, userId from syn18513076")$filepath %>%
  read_csv()

leaderboard <- read_csv(syn$get("syn18520916")$path) %>% full_join(fv)

get_path <- function(id){
  syn$get(id)$path
}

gold <- read_csv(syn$get("syn18421225")$path)
```

```{r echo=FALSE}
get_user_or_team_names <- function(id){
 name <- try(syn$getTeam(id)$name, silent = T) ##try to get the team name from id 
 if(class(name)=='try-error'){ ##if it is not a team, will return error, so then try to get user profile
 try({
   prof <- syn$getUserProfile(id = id) ##get first and last name
   fn <- prof$firstName
   ln <- prof$lastName
   if(is.null(fn) | is.null(ln)){
     un <- prof$userName
     return(un) 
   }else if(fn == "" | ln == ""){ ##if empty, get username instead
     un <- prof$userName
     return(un)
   }else{
     return(paste(fn, ln))
   }
   })
   }else{
     return(name)
   }
}

names<- sapply(unique(leaderboard$submitterId), get_user_or_team_names)
names(names) <- unique(leaderboard$submitterId)
names_df <- as.data.frame(names) %>% 
  magrittr::set_colnames("participant") %>% 
  rownames_to_column("submitterId") %>% 
  mutate(submitterId= as.numeric(submitterId)) %>% 
  mutate(participant = as.character(participant))

names_df$participant[names_df$participant=="瑞 张"] <- syn$getUserProfile(id = names_df$submitterId[names_df$participant=="瑞 张"])$userName

names_df$participant[names_df$submitterId=="3383197"] <- "Prospectors"


submitted_docker_and_methods <- c("3379336", #baseline
                                  "3361565", #n121
                                  "3378762", #KinaseHunter
                                  "3379046", #Kermit-lab
                                  "3377646", #QED
                                  "3379110", #KKT
                                  "3379198", #let data talk
                                  "3383197", #davor orsolic/prospectors
                                  "3380669", #boun
                                  "3380843", #druginaselearning
                                  "2223251", #mehmet tan, oselot
                                  "3361766", #thinng
                                  "3384382", #dmisdk
                                  "3371544", #metu_embl
                                  "3324878", #hulab
                                  "3337435", #aydin
                                  "3332429", #ai winter is coming
                                  "3380028", #ml med
                                  "3380320", #amsterdamumc-ku
                                  "3343575") #MCIC


```

### SC1 - SPEARMAN 

```{r echo=FALSE, message=FALSE, warning=FALSE}
####SPEARMAN ANALYSIS

best_spearman_by_team <- leaderboard %>%
  group_by(submitterId) %>%
  top_n(1, spearman) %>%
  ungroup() %>%
  arrange(-spearman) %>% 
  left_join(names_df)

paths <- map_chr(best_spearman_by_team$id, get_path)

best_path <- paths[1]
other_paths <- paths[-1]

results_spearman <- finalScoring(predictions = other_paths,
             predictionColname = 'pKd_[M]_pred',
             predictionIds = best_spearman_by_team$objectId[-1],
             goldStandard = gold,
             goldStandardColname = 'pKd_[M]',
             bestPrediction = best_path,
             bestPredictionId = best_spearman_by_team$objectId[1],
             keyColumns = colnames(gold)[1:6],
             doParallel = TRUE,
             scoreFun = spearman_py)

tidy_bayes <- tibble('objectId' = colnames(results_spearman$bootstrappedScores),
                         'bayes' = results_spearman$bayes)


tidy_res <- results_spearman$bootstrappedScores %>%
  as.data.frame %>%
  tidyr::gather(objectId, bootstrappedScore) %>%
  left_join(tidy_bayes) %>%
  left_join(leaderboard %>% mutate(objectId = as.character(objectId))) %>% 
  left_join(names_df) %>% 
  mutate(dq = case_when(submitterId %in% submitted_docker_and_methods ~ "",
                   !submitterId %in% submitted_docker_and_methods ~ "dq"))

dqed <- tidy_res %>% group_by(participant, dq) %>% 
  summarize(med = median(bootstrappedScore)) %>% 
  arrange(med)
dqed_vec  <- ifelse(dqed$dq=="dq", "azure4", "black")
dqed_vec <- dqed_vec[order(dqed$med)]

ggplot(data = tidy_res) +
  geom_boxplot(aes(x = reorder(participant, bootstrappedScore, fun = median), y = bootstrappedScore,
                    color = cut(bayes, c(-2,-1,0, 3, 5,Inf), include.lowest = T, ordered_result = T, right = T)), outlier.shape = NA)+
  coord_flip() +
  scale_color_manual(values = c("#AA1155","#F8333C","#FCAB10","#2B9EB3"),
                     name = "Bayes Factor",
                     labels = c("Reference", "<3","3-5",">5")) +
  labs(x = "Team/Participant", y = "Bootstrapped Spearman") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6, color = dqed_vec))

```

### SC2 - RMSE

```{r echo=FALSE, message=FALSE, warning=FALSE}

####RMSE ANALYSIS

best_rmse_by_team <- leaderboard %>%
  group_by(submitterId) %>%
  top_n(1, -rmse) %>%
  ungroup() %>%
  arrange(rmse) %>% 
  left_join(names_df)

paths <- map_chr(best_rmse_by_team$id, get_path)

best_path <- paths[2]
other_paths <- paths[-2]

results_rmse <- finalScoring(predictions = other_paths,
                        predictionColname = 'pKd_[M]_pred',
                        predictionIds = best_rmse_by_team$objectId[-2],
                        goldStandard = gold,
                        goldStandardColname = 'pKd_[M]',
                        bestPrediction = best_path,
                        bestPredictionId = best_rmse_by_team$objectId[2],
                        keyColumns = colnames(gold)[1:6],
                        doParallel = TRUE,
                        scoreFun = rmse_py,
                        largerIsBetter = F)

# recalc <- computeBayesFactorWhereRefIsNotBest(results_rmse$bootstrappedScores, 2, invertBayes = F)

tidy_bayes <- tibble('objectId' = colnames(results_rmse$bootstrappedScores),
                     'bayes' = results_rmse$bayes)


tidy_res <- results_rmse$bootstrappedScores %>%
  as.data.frame %>%
  tidyr::gather(objectId, bootstrappedScore) %>%
  left_join(tidy_bayes) %>%
  left_join(leaderboard %>% mutate(objectId = as.character(objectId))) %>% 
  left_join(names_df) %>% 
  mutate(dq = case_when(submitterId %in% submitted_docker_and_methods ~ "",
                   !submitterId %in% submitted_docker_and_methods ~ "dq"))

dqed <- tidy_res %>% group_by(participant, dq) %>% 
  summarize(med = median(bootstrappedScore)) %>% 
  arrange(med)
dqed_vec  <- ifelse(dqed$dq=="dq", "azure4", "black")
dqed_vec <- dqed_vec[order(-dqed$med)]

ggplot(data = tidy_res) +
  geom_boxplot(aes(x = reorder(participant, -bootstrappedScore, fun = median), y = log10(bootstrappedScore),
                    color = cut(bayes, c(-2,-1,0, 3, 5,Inf), include.lowest = T, ordered_result = T, right = T)), outlier.shape = NA)+
  coord_flip() +
  scale_color_manual(values = c("#AA1155","#F8333C","#FCAB10","#2B9EB3"),
                     name = "Bayes Factor",
                     labels = c("Reference", "<3","3-5",">5")) +
  labs(x = "Team/Participant", y = "log10(Bootstrapped RMSE)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6, color = dqed_vec))
```

Now, let's reanalyze the above plots, but using the baseline as the reference for the Bayes factor calculation.

### SC1 - SPEARMAN 

```{r echo=FALSE, message=FALSE, warning=FALSE}
####SPEARMAN ANALYSIS

idx_plos <- match("9685944",colnames(results_spearman$bootstrappedScores))

results_spearman$recalculate_bayes <- computeBayesFactorWhereRefIsNotBest(results_spearman$bootstrappedScores, idx_plos ,invertBayes = F)

tidy_bayes <- tibble('objectId' = colnames(results_spearman$bootstrappedScores),
                         'bayes' = results_spearman$recalculate_bayes)


tidy_res <- results_spearman$bootstrappedScores %>%
  as.data.frame %>%
  tidyr::gather(objectId, bootstrappedScore) %>%
  left_join(tidy_bayes) %>%
  left_join(leaderboard %>% mutate(objectId = as.character(objectId))) %>% 
  left_join(names_df) %>% 
  mutate(dq = case_when(submitterId %in% submitted_docker_and_methods ~ "",
                   !submitterId %in% submitted_docker_and_methods ~ "dq"))

dqed <- tidy_res %>% group_by(participant, dq) %>% 
  summarize(med = median(bootstrappedScore)) %>% 
  arrange(med)
dqed_vec  <- ifelse(dqed$dq=="dq", "azure4", "black")
dqed_vec <- dqed_vec[order(dqed$med)]

ggplot(data = tidy_res) +
  geom_boxplot(aes(x = reorder(participant, bootstrappedScore, fun = median), y = bootstrappedScore,
                    color = cut(bayes, c(-2,-1,0, 3, 5,Inf), include.lowest = T, ordered_result = T, right = T)), outlier.shape = NA)+
  coord_flip() +
  scale_color_manual(values = c("#AA1155","#F8333C","#FCAB10","#2B9EB3"),
                     name = "Bayes Factor",
                     labels = c("Reference", "<3","3-5",">5")) +
  labs(x = "Team/Participant", y = "Bootstrapped Spearman") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6, color = dqed_vec))
```

### SC2 - RMSE

```{r echo=FALSE, message=FALSE, warning=FALSE}

####RMSE ANALYSIS
idx_plos <- match("9685944",colnames(results_rmse$bootstrappedScores))

results_rmse$recalculate_bayes <- computeBayesFactorWhereRefIsNotBest(results_rmse$bootstrappedScores, idx_plos ,invertBayes = F)

tidy_bayes <- tibble('objectId' = colnames(results_rmse$bootstrappedScores),
                         'bayes' = results_rmse$recalculate_bayes)


tidy_res <- results_rmse$bootstrappedScores %>%
  as.data.frame %>%
  tidyr::gather(objectId, bootstrappedScore) %>%
  left_join(tidy_bayes) %>%
  left_join(leaderboard %>% mutate(objectId = as.character(objectId))) %>% 
  left_join(names_df) %>% 
  mutate(dq = case_when(submitterId %in% submitted_docker_and_methods ~ "",
                   !submitterId %in% submitted_docker_and_methods ~ "dq"))

dqed <- tidy_res %>% group_by(participant, dq) %>% 
  summarize(med = median(bootstrappedScore)) %>% 
  arrange(med)
dqed_vec  <- ifelse(dqed$dq=="dq", "azure4", "black")
dqed_vec <- dqed_vec[order(-dqed$med)]

ggplot(data = tidy_res) +
  geom_boxplot(aes(x = reorder(participant, -bootstrappedScore, fun = median), y = log10(bootstrappedScore),
                    color = cut(bayes, c(-2,-1,0, 3, 5,Inf), include.lowest = T, ordered_result = T, right = T)), outlier.shape = NA)+
  coord_flip() +
  scale_color_manual(values = c("#AA1155","#F8333C","#FCAB10","#2B9EB3"),
                     name = "Bayes Factor",
                     labels = c("Reference", "<3","3-5",">5")) +
  labs(x = "Team/Participant", y = "log10(Bootstrapped RMSE)") +
  theme_bw() +
  theme(axis.text.y = element_text(size = 6, color = dqed_vec))

```

```{r}
sessionInfo()

```