---
title: "Compare R2 predictions to random and theoretical upper bound"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

###Setup

```{r}
library(reticulate)
library(tidyverse)

set.seed(13048345)

use_python("/usr/local/bin/python2")
synapse <- import("synapseclient")
syn <- synapse$Synapse()
synutils <- synapse$utils
syn$login()
source_python('https://raw.githubusercontent.com/Sage-Bionetworks/IDG-DREAM-Drug-Kinase-Challenge/master/round1b/score/bin/evaluation_metrics_python2.py')

spearman_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  spearman(gold_py, pred_py)
}

rmse_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  rmse(gold_py, pred_py)
}

set.seed(90210)

# template <- syn$get("syn16809885")$path %>% read_csv
# dtc <- syn$get("syn18061798")$path %>% read_csv
# filt <- dtc %>% filter(standard_type == "PKD", standard_relation == "=")

gold <- syn$get("syn18421225")$path %>% read_csv


fv <- syn$tableQuery("select id, submissionId AS objectId, teamId, userId from syn18513076")$filepath %>%
  read_csv()

leaderboard <- read_csv(syn$get("syn18520916")$path) %>% full_join(fv) 

get_user_or_team_names <- function(id){
 name <- try(syn$getTeam(id)$name, silent = T) ##try to get the team name from id 
 if(class(name)=='try-error'){ ##if it is not a team, will return error, so then try to get user profile
 try({
   prof <- syn$getUserProfile(id = id) ##get first and last name
   fn <- prof$firstName
   ln <- prof$lastName
   if(is.null(fn) | is.null(ln)){
     un <- prof$userName
     return(un) 
   }else if(fn == "" | ln == ""){ ##if empty, get username instead
     un <- prof$userName
     return(un)
   }else{
     return(paste(fn, ln))
   }
   })
   }else{
     return(name)
   }
}

names<- sapply(unique(leaderboard$submitterId), get_user_or_team_names)
names(names) <- unique(leaderboard$submitterId)
names_df <- as.data.frame(names) %>% 
  magrittr::set_colnames("participant") %>% 
  rownames_to_column("submitterId") %>% 
  mutate(submitterId= as.numeric(submitterId)) %>% 
  mutate(participant = as.character(participant))

names_df$participant[names_df$participant=="瑞 张"] <- syn$getUserProfile(id = names_df$submitterId[names_df$participant=="瑞 张"])$userName

names_df$participant[names_df$submitterId=="3383197"] <- "Prospectors"


```

### Define best predictions for R2 

For each metric.
We will look at all submissions.

```{r echo=FALSE, message=FALSE, warning=FALSE}
# 
# best_spearman_by_team <- leaderboard %>%
#   group_by(submitterId) %>%
#   top_n(1, spearman) %>%
#   ungroup() %>%
#   arrange(-spearman) %>% 
#   left_join(names_df) 
# 
# best_rmse_by_team <- leaderboard %>%
#   group_by(submitterId) %>%
#   top_n(1, rmse) %>%
#   ungroup() %>%
#   arrange(-rmse) %>% 
#   left_join(names_df) 


spearman_by_team <- leaderboard %>%
  arrange(-spearman) %>%
  left_join(names_df)

rmse_by_team <- leaderboard %>%
  arrange(-rmse) %>%
  left_join(names_df)

```

### Compare predictions to random 

Random prediction distribution made by sampling the test data Kds 10000 times and assigning to the test comparisons, scoring each sample, and plotting the distributions of the scores. 

We can then plot the actual scores on the x-axis to compare which fall above (spearman) or below (rmse) the distribution, suggesting they did better than random. 

```{r}
vec_sp <- sapply(1:10000, function(x){
  foo<- sample(gold$`pKd_[M]`, nrow(gold))
  spearman_py(foo, gold$`pKd_[M]`)
}) %>% as.data.frame() %>% set_names(c('spearman'))

vec_rm <- sapply(1:10000, function(x){
  foo<- sample(gold$`pKd_[M]`, nrow(gold))
  rmse_py(foo, gold$`pKd_[M]`) 
}) %>% as.data.frame() %>% set_names(c('rmse'))

p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#E5C8AC" ,fill = "#E5C8AC", alpha = .5 , bw = "sj") +
  geom_point(data = spearman_by_team, aes(x=spearman),  color =  "#DB5700",  y = 0) +
  theme_bw() + 
  annotate(geom = "text", x = 0.4, y = 6,
             label = paste0(signif(length(spearman_by_team$spearman[spearman_by_team$spearman>max(vec_sp)])/length(spearman_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
             size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density")
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = log(rmse)), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .5, bw = "sj") +
  geom_point(data = rmse_by_team %>% filter(rmse < 5), aes(x=log(rmse)), color = "#046C9A", y = 0) +
  theme_bw() + 
  annotate(geom = "text", x = 0, y = 10,
             label = paste0(signif(length(rmse_by_team$rmse[rmse_by_team$rmse<min(vec_rm)])/length(rmse_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
             size = 5) +
  xlim(c(-.25,0.75)) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "log10(RMSE)", y= "Density")


gridExtra::grid.arrange(p1,p2)

```

We can do a similar analysis to assess the theoretical upper bound of prediction scoring using some replicate data to assess the variability of Kd measurements. 


```{r}
axitinib_reps <- syn$get("syn20350570")$path
load(axitinib_reps)

axitinib <- df
#   set_names(c("public_pKd", "pKd_[M]")) %>% 
#   right_join(gold)
# 
# axitinib_join <- gold %>% 
#   filter()

vec_sp_max <- sapply(1:10000, function(x){
  foo<- axitinib %>% sample_n(nrow(axitinib), replace = T)
  spearman_py(foo$public_pKd, foo$new_pKd)
}) %>% as.data.frame() %>% set_names(c('spearman'))

vec_rm_max <- sapply(1:10000, function(x){
  foo<- axitinib %>% sample_n(nrow(axitinib), replace = T)
  rmse_py(foo$public_pKd, foo$new_pKd)
}) %>% as.data.frame() %>% set_names(c('rmse'))


p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#E5C8AC" ,fill = "#E5C8AC", alpha = .2 , bw = "sj") +
  geom_density(data = vec_sp_max, aes(x = spearman), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .9, bw = "sj") +
  geom_point(data = spearman_by_team, aes(x=spearman),  color =  "#DB5700",  y = 0) +
  theme_bw() + 
  # annotate(geom = "text", x = 0.4, y = 6,
  #            label = paste0(signif(length(spearman_by_team$spearman[spearman_by_team$spearman>max(vec_sp)])/length(spearman_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
  #            size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density")
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .2, bw = "sj") +
  geom_density(data = vec_rm_max, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .9, bw = "sj") +
  geom_point(data = rmse_by_team %>% filter(rmse < 5), aes(x=rmse), color = "#046C9A", y = 0) +
  theme_bw() + 
  # annotate(geom = "text", x = 0, y = 10,
  #            label = paste0(signif(length(rmse_by_team$rmse[rmse_by_team$rmse<min(vec_rm)])/length(rmse_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
  #            size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "RMSE", y= "Density")


gridExtra::grid.arrange(p1,p2)

```

The distribution of the "upper bound" is pretty wide. I don't think this is a great representation of experimental variability. Instead, we chose to look at all replicates in DTC for all drugs tested in R2 (different pairs though, since it's already public data). Balaguru provided this data.

After discussion in the group, we decided to filter out the replicates where pKd=5 so that they only represent ~25% of the final dataset, which is closer in proportion to the test R2 datasets where pKd=5. This is to get a better estimate of the max upper bound, particularly RMSE, which is substantially affected by a large proportion of pKd =5 values.

```{r}
replicates <- syn$get('syn20448785')$path %>%
  read_tsv %>% 
  spread(Pubmed_id, pKd) 

replicates_5 <- filter(replicates, `15711537` == 5 & `22037378` == 5) %>% 
  sample_n(104) 

replicates_subset <- replicates %>% 
  filter(`15711537` > 5 | `22037378` > 5) %>% 
  bind_rows(replicates_5) 

vec_sp_max <- sapply(1:10000, function(x){
  foo<- replicates_subset %>% sample_n(nrow(gold), replace = T)
  spearman_py(foo$`15711537`, foo$`22037378`)
}) %>% as.data.frame() %>% set_names(c('spearman'))

vec_rm_max <- sapply(1:10000, function(x){
  foo<- replicates_subset %>% sample_n(nrow(gold), replace = T)
  rmse_py(foo$`15711537`, foo$`22037378`)
}) %>% as.data.frame() %>% set_names(c('rmse'))


p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#ABDDDE" ,fill = "#ABDDDE", alpha = .2 , bw = "sj") +
  geom_density(data = vec_sp_max, aes(x = spearman), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .9, bw = "sj") +
  geom_point(data = spearman_by_team %>% 
               filter(spearman>max(vec_sp)),
             aes(x=spearman),  color =  "#046C9A",  y = 0) +
    geom_point(data = spearman_by_team %>% 
               filter(spearman<=max(vec_sp)),
             aes(x=spearman),  color =  "#046C9A",  y = 0, shape = 1, alpha = 0.7) +
  theme_bw() + 
  # annotate(geom = "text", x = 0.4, y = 6,
  #            label = paste0(signif(length(spearman_by_team$spearman[spearman_by_team$spearman>max(vec_sp)])/length(spearman_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
  #             size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density") +
 xlim(c(-.25, 1))
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .2, bw = "sj") +
  geom_density(data = vec_rm_max, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .9, bw = "sj") +
  geom_point(data = rmse_by_team %>% 
               filter(rmse<min(vec_rm) & rmse < 2),
             aes(x=rmse),  color =  "#046C9A",  y = 0) +
    geom_point(data = rmse_by_team %>% 
               filter(rmse>=min(vec_rm)),
             aes(x=rmse),  color =  "#046C9A",  y = 0, shape = 1, alpha = 0.7) +  theme_bw() + 
   # annotate(geom = "text", x = 0, y = 10,
   #            label = paste0(signif(length(rmse_by_team$rmse[rmse_by_team$rmse<min(vec_rm)])/length(rmse_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
              # size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "RMSE", y= "Density") +
  xlim(c(0, 1.75))


gridExtra::grid.arrange(p1,p2)


print(spearman_by_team$objectId[spearman_by_team$spearman>max(vec_sp)]) 
print(rmse_by_team$objectId[rmse_by_team$rmse<min(vec_rm)]) 


```

Same fig as above but with theme for paper. Doesn't render well in markdown. 

```{r}
library(Cairo)

p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#ABDDDE" ,fill = "#ABDDDE", alpha = .2 , bw = "sj") +
  geom_density(data = vec_sp_max, aes(x = spearman), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .9, bw = "sj") +
  geom_point(data = spearman_by_team %>% 
               filter(spearman>max(vec_sp)),
             aes(x=spearman),  color =  "#046C9A",  y = 0) +
    geom_point(data = spearman_by_team %>% 
               filter(spearman<=max(vec_sp)),
             aes(x=spearman),  color =  "#046C9A",  y = 0, shape = 1, alpha = 0.7) +
  theme_bw() + 
  # annotate(geom = "text", x = 0.4, y = 6,
  #            label = paste0(signif(length(spearman_by_team$spearman[spearman_by_team$spearman>max(vec_sp)])/length(spearman_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
  #             size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density") +
  xlim(c(-.25, 1)) +
  ylim(c(0, 30)) +
  theme(axis.text = element_text(size = 40), # face = "bold"
        axis.title = element_text(size = 40),
        axis.title.x = element_blank()
  ) + 
  theme(legend.position = "none") + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))


ggsave("figure_3_spearman_r2.pdf",  device = cairo_pdf,
              width = 11.69, height = 4.135, units = "in")
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .2, bw = "sj") +
  geom_density(data = vec_rm_max, aes(x = rmse), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .9, bw = "sj") +
  geom_point(data = rmse_by_team %>% 
               filter(rmse<min(vec_rm) & rmse < 2),
             aes(x=rmse),  color =  "#046C9A",  y = 0) +
    geom_point(data = rmse_by_team %>% 
               filter(rmse>=min(vec_rm)),
             aes(x=rmse),  color =  "#046C9A",  y = 0, shape = 1, alpha = 0.7) +  theme_bw() + 
   # annotate(geom = "text", x = 0, y = 10,
   #            label = paste0(signif(length(rmse_by_team$rmse[rmse_by_team$rmse<min(vec_rm)])/length(rmse_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
              # size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "RMSE", y= "Density") +
  xlim(c(0, 2)) +
  ylim(c(0, 18)) +
  theme(axis.text = element_text(size = 40), # face = "bold"
        axis.title = element_text(size = 40),
        axis.title.x = element_blank()
  ) + 
  theme(legend.position = "none") + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))


ggsave("figure_3_rmse_r2.pdf",  device = cairo_pdf,
              width = 11.69, height = 4.135, units = "in")

```

Here's some other test analyses, iteratively scoring team QED submission in order of absolute error of predictions to assess when score falls below max theoretical threshold: 

```{r}

qed <- syn$get('syn18513191')$path %>% read_csv

data <- gold %>% 
  left_join(qed) %>% 
  mutate(ae = abs(`pKd_[M]`-`pKd_[M]_pred`))  %>% 
  arrange(ae)

bar <- c()

n_pair_test <- c(40,80,120,160,200,240,280,320,340,380,394)
for(i in n_pair_test){
  foo <- top_n(data, i, -ae) %>% 
    add_column(n_pairs = i)
  
  bar <- bind_rows(bar,foo)
}

ggplot(bar)+
  geom_boxplot(aes(x=as.factor(n_pairs),y=ae)) +
  theme_bw() +
  labs(x = "Number of Pairs", y = "Absolute Error")


iter_score <- function(data,n_iter){
  foo <- slice(data,1:n_iter)
  spearman_py(foo$`pKd_[M]`, foo$`pKd_[M]_pred`)
}

bar <- c()

for(i in 1:nrow(data)){
  bar<- c(bar,iter_score(data, i))
}

data$iter_score <- bar

ggplot(data)+
  geom_point(aes(x=ae,y=iter_score, color = iter_score > min(vec_sp_max))) +
  geom_hline(yintercept = min(vec_sp_max), color = "#DB5700") +
  theme_bw() +
  labs(x = "Absolute Error", y = "Spearman Correlation")


data_summary <- data %>% 
  group_by(Compound_Name) %>% 
  summarize(mean_ae = mean(ae))

ggplot(data)+
  geom_boxplot(aes(x=reorder(Compound_Name, ae, FUN = median),, y = ae)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

ggplot(data %>% group_by(UniProt_Id) %>% add_count(UniProt_Id) %>% filter(n>3))+
  geom_boxplot(aes(x=reorder(Entrez_Gene_Symbol, ae, FUN = median),, y = ae)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))

vec <- data$iter_score > min(vec_sp_max)
length(vec[vec==TRUE]) ##length of the number of pairs when combined is above min theoretical max


```

