---
title: "Compare R1 predictions to random and theoretical upper bound"
output:
  html_document:
    df_print: paged
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

###Setup

```{r}
library(reticulate)
library(tidyverse)

set.seed(13048345)

use_python("/usr/local/bin/python2")
synapse <- import("synapseclient")
syn <- synapse$Synapse()
synutils <- synapse$utils
syn$login()
source_python('https://raw.githubusercontent.com/Sage-Bionetworks/IDG-DREAM-Drug-Kinase-Challenge/master/round1b/score/bin/evaluation_metrics_python2.py')

spearman_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  spearman(gold_py, pred_py)
}

rmse_py <- function(gold, pred){
  gold_py <- gold %>% np_array()
  pred_py <- pred %>% np_array()
  rmse(gold_py, pred_py)
}

set.seed(90210)

# template <- syn$get("syn16809885")$path %>% read_csv
# dtc <- syn$get("syn18061798")$path %>% read_csv
# filt <- dtc %>% filter(standard_type == "PKD", standard_relation == "=")

gold <- syn$get("syn16809884")$path %>% read_csv


fv <- syn$tableQuery("select id, submissionId AS objectId, teamId, userId from syn17051994", includeRowIdAndRowVersion = F)$filepath %>%
  read_csv

sub_id_map <- syn$get("syn17054264")$path %>% read_csv %>% select(objectId, submitterId)

leaderboard <- syn$tableQuery("SELECT objectId, rmse, spearman FROM syn17054253", includeRowIdAndRowVersion = F)$filepath %>% 
  read_csv%>% full_join(fv) %>% left_join(sub_id_map)

get_user_or_team_names <- function(id){
 name <- try(syn$getTeam(id)$name, silent = T) ##try to get the team name from id 
 if(class(name)=='try-error'){ ##if it is not a team, will return error, so then try to get user profile
 try({
   prof <- syn$getUserProfile(id = id) ##get first and last name
   fn <- prof$firstName
   ln <- prof$lastName
   if(is.null(fn) | is.null(ln)){
     un <- prof$userName
     return(un) 
   }else if(fn == "" | ln == ""){ ##if empty, get username instead
     un <- prof$userName
     return(un)
   }else{
     return(paste(fn, ln))
   }
   })
   }else{
     return(name)
   }
}

names<- sapply(unique(leaderboard$submitterId), get_user_or_team_names)
names(names) <- unique(leaderboard$submitterId)
names_df <- as.data.frame(names) %>% 
  magrittr::set_colnames("participant") %>% 
  rownames_to_column("submitterId") %>% 
  mutate(submitterId= as.numeric(submitterId)) %>% 
  mutate(participant = as.character(participant))

names_df$participant[names_df$submitterId=="3383197"] <- "Prospectors"


```

### Define best predictions for R1

For each metric.
Note, we'll not actually do this, instead we will look at all submissions

```{r echo=FALSE, message=FALSE, warning=FALSE}
# 
# best_spearman_by_team <- leaderboard %>%
#   group_by(submitterId) %>%
#   top_n(1, spearman) %>%
#   ungroup() %>%
#   arrange(-spearman) %>% 
#   left_join(names_df) 
# 
# best_rmse_by_team <- leaderboard %>%
#   group_by(submitterId) %>%
#   top_n(1, rmse) %>%
#   ungroup() %>%
#   arrange(-rmse) %>% 
#   left_join(names_df) 


spearman_by_team <- leaderboard %>%
  arrange(-spearman) %>%
  left_join(names_df)

rmse_by_team <- leaderboard %>%
  arrange(-rmse) %>%
  left_join(names_df)

```

### Compare predictions to random 

Random prediction distribution made by sampling the test data Kds 10000 times and assigning to the test comparisons, scoring each sample, and plotting the distributions of the scores. 

We can then plot the actual scores on the x-axis to compare which fall above (spearman) or below (rmse) the distribution, suggesting they did better than random. 

```{r}
vec_sp <- sapply(1:10000, function(x){
  foo<- sample(gold$`pKd_[M]`, nrow(gold))
  spearman_py(foo, gold$`pKd_[M]`)
}) %>% as.data.frame() %>% set_names(c('spearman'))

vec_rm <- sapply(1:10000, function(x){
  foo<- sample(gold$`pKd_[M]`, nrow(gold))
  rmse_py(foo, gold$`pKd_[M]`) 
}) %>% as.data.frame() %>% set_names(c('rmse'))

p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#E5C8AC" ,fill = "#E5C8AC", alpha = .5 , bw = "sj") +
  geom_point(data = spearman_by_team, aes(x=spearman),  color =  "#DB5700",  y = 0) +
  theme_bw() + 
  annotate(geom = "text", x = 0.4, y = 6,
             label = paste0(signif(length(spearman_by_team$spearman[spearman_by_team$spearman>max(vec_sp)])/length(spearman_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
             size = 5) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density")
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = log(rmse)), color =  "#ABDDDE",  fill = "#ABDDDE", alpha = .5, bw = "sj") +
  geom_point(data = rmse_by_team %>% filter(rmse < 5), aes(x=log(rmse)), color = "#046C9A", y = 0) +
  theme_bw() + 
  annotate(geom = "text", x = 0, y = 10,
             label = paste0(signif(length(rmse_by_team$rmse[rmse_by_team$rmse<min(vec_rm)])/length(rmse_by_team$rmse)*100, 2), "% of submissions \nbetter than random") ,
             size = 5) +
  xlim(c(-.25,0.75)) +
  theme(axis.text = element_text(size = 12)) +
  labs(x = "log10(RMSE)", y= "Density")


gridExtra::grid.arrange(p1,p2)

```

For the upper bound, we are looking at all replicates in DTC for all drugs tested in R2 (different pairs though, since it's already public data). Balaguru provided this data.

After discussion in the group, we decided to filter out the replicates where pKd=5 so that they only represent ~35% of the final dataset, which is closer in proportion to the test R1 ataseT where pKd=5. This is to get a better estimate of the max upper bound, particularly RMSE, which is substantially affected by a large proportion of pKd =5 values.

```{r}
replicates <- syn$get('syn20448785')$path %>% ##r1 kinases and compounds from r2 data
  read_tsv %>% 
  spread(Pubmed_id, pKd) %>% 
  filter(Compound_InchiKeys %in% gold$Compound_InchiKeys | UniProt_Id %in% gold$UniProt_Id)

replicates_all_r1 <- syn$get("syn20487871")$path %>% ##remaining DTC pairs provided seperately
  read_tsv %>% 
  spread(Pubmed_id, pKd) %>% 
  bind_rows(replicates) %>% 
  distinct()

replicates_5 <- filter(replicates_all_r1, `15711537` == 5 & `22037378` == 5) %>% 
  sample_n(111) 

replicates_subset <- replicates %>% 
  filter(`15711537` > 5 | `22037378` > 5) %>% 
  bind_rows(replicates_5) 


vec_sp_max <- sapply(1:10000, function(x){
  foo<- replicates_subset %>% sample_n(nrow(gold), replace = T)
  spearman_py(foo$`15711537`, foo$`22037378`)
}) %>% as.data.frame() %>% set_names(c('spearman'))

vec_rm_max <- sapply(1:10000, function(x){
  foo<- replicates_subset %>% sample_n(nrow(gold), replace = T)
  rmse_py(foo$`15711537`, foo$`22037378`)
}) %>% as.data.frame() %>% set_names(c('rmse'))


p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#E5C8AC" ,fill = "#E5C8AC", alpha = .2 , bw = "sj") +
  geom_density(data = vec_sp_max, aes(x = spearman), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .9, bw = "sj") +
  geom_point(data = spearman_by_team %>% 
               filter(spearman>max(vec_sp)),
             aes(x=spearman),  color =  "#DB5700",  y = 0) +
  geom_point(data = spearman_by_team %>% 
               filter(spearman<=max(vec_sp)),
             aes(x=spearman),  color =  "#DB5700",  y = 0, shape = 1, alpha = 0.7) +
  theme_bw() + 
  theme(axis.text = element_text(size = 12)) +
  labs(x = "Spearman Correlation", y= "Density") +
 xlim(c(-.25, 1))
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = rmse), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .2, bw = "sj") +
  geom_density(data = vec_rm_max, aes(x = rmse), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .9, bw = "sj") +
  geom_point(data = rmse_by_team %>% 
               filter(rmse<min(vec_rm) & rmse < 2),
             aes(x=rmse),  color =  "#DB5700",  y = 0) +
  geom_point(data = rmse_by_team %>% 
               filter(rmse>=min(vec_rm)),
             aes(x=rmse),  color =  "#DB5700",  y = 0, shape = 1, alpha = 0.7) +  theme_bw() + 
  theme_bw() + 
  theme(axis.text = element_text(size = 12)) +
  labs(x = "RMSE", y= "Density") +
  xlim(c(0, 2))


gridExtra::grid.arrange(p1,p2)


print(spearman_by_team$objectId[spearman_by_team$spearman>max(vec_sp)]) 
print(rmse_by_team$objectId[rmse_by_team$rmse<min(vec_rm)]) 


```

Same fig as above but with theme for paper. Doesn't render well in markdown. 

```{r}
library(Cairo)

p1 <- ggplot() +
  geom_density(data = vec_sp, aes(x = spearman),color = "#E5C8AC" ,fill = "#E5C8AC", alpha = .2 , bw = "sj") +
  geom_density(data = vec_sp_max, aes(x = spearman), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .9, bw = "sj") +
  geom_point(data = spearman_by_team %>% 
               filter(spearman>max(vec_sp)),
             aes(x=spearman),  color =  "#DB5700",  y = 0) +
  geom_point(data = spearman_by_team %>% 
               filter(spearman<=max(vec_sp)),
             aes(x=spearman),  color =  "#DB5700",  y = 0, shape = 1, alpha = 0.7) +
  theme_bw() + 
  labs(x = "Spearman Correlation", y= "Density") +
  xlim(c(-.25, 1)) +
  ylim(c(0, 30)) +
  theme(axis.text = element_text(size = 40), # face = "bold"
        axis.title = element_text(size = 40),
        axis.title.x = element_blank()
  ) + 
  theme(legend.position = "none") + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))

ggsave("figure_3_spearman_r1.pdf",  device = cairo_pdf,
              width = 11.69, height = 4.135, units = "in")
  
p2 <- ggplot() +
  geom_density(data = vec_rm, aes(x = rmse), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .2, bw = "sj") +
  geom_density(data = vec_rm_max, aes(x = rmse), color =  "#E5C8AC",  fill = "#E5C8AC", alpha = .9, bw = "sj") +
  geom_point(data = rmse_by_team %>% 
               filter(rmse<min(vec_rm) & rmse < 2),
             aes(x=rmse),  color =  "#DB5700",  y = 0) +
  geom_point(data = rmse_by_team %>% 
               filter(rmse>=min(vec_rm)),
             aes(x=rmse),  color =  "#DB5700",  y = 0, shape = 1, alpha = 0.7) +  theme_bw() + 
  theme_bw() + 
  labs(x = "RMSE", y= "Density") +
  xlim(c(0, 2)) +
  ylim(c(0, 18)) +
  theme(axis.text = element_text(size = 40), # face = "bold"
        axis.title = element_text(size = 40),
        axis.title.x = element_blank()
  ) + 
  theme(legend.position = "none") + 
  theme(plot.margin = unit(c(1,0.5,0.5,0.5), "cm"))

ggsave("figure_3_rmse_r1.pdf",  device = cairo_pdf,
              width = 11.69, height = 4.135, units = "in")

p1
p2

```

```
